---
title: "AIDS Drug"
author: "Tyler Hamblin"
date: "21 September 2023"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

$H_0$: There is no significant change in infection rate when a treatment drug is used. 

$H_a$: There is a significant change in infection rate when a treatment drug is used. 

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

The Dplyr tool "filter" can be used to separate out a subset of data from the original set provided to us by the question. After collecting only individuals that are drug users we can use CandP helper files to give us graphical, numerical, and inferential data to prove or reject the null hypothesis. 



```{r include=FALSE,echo=FALSE}
drug_users <- filter(triUmp, DrugUser == "yes")

drug_users
```

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=drug_users)
barchartGC(~infected + treatment,data=drug_users, type="percent")

```

This code has created two bar charts in which the first shows the total number of people involved in this statistical analysis. 

Chart 1 suggests that most of the people who are a drug users contract AIDS less frequently than people in the placebo category. However, this does not give the clearest picture of the effectiveness of the drug. 

Chart 2 expresses the data as a percentage and therefore it becomes clear that the percentage of people contracting AIDS on the drug (~25%) when compared to the placebo (~75%) is significantly fewer. 


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~infected + treatment, data=drug_users)
rowPerc(table1)
colPerc(table1)
```

These two charts break down a numerical version of the previous bar charts and the rate at which test subjects contract AIDS based on their treatment group. As seen in both the top and bottom, those that tested positive for AIDS were about three times as likely to fall into the placebo category than the drug user category. This proves that the treatment method is effective at preventing the contraction of AIDS but it does not have a 100% success rate. 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

There are three inferential tests that we ran in this section with two being Chi-Squared Tests and the other being a Fisher Exact Test. The Chi-Squared Tests give us a very minute p-value of about 0.0000001242 which is far lower than the typical 0.05 that informally suggests the cutoff for a significance of data. The Chi-Squared value of -28 is much larger than the critical value for degrees of freedom, which in this data set is (1). All three of these tests suggest that there is a statistical significance in results. 


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Even just in the graphical and numerical results we notice that the placebo group tends to contract the disease at a far greater rate than those who were treated with the drug. Along with the tests below in the inferential section it confirms that there is a statistical significance in this group of data. This in turn proves that there is a negative relationship between the drug users and the disease frequency. Since there is a large amount of people that were surveyed for this study it can be assumed that random variations played a part in the results but were not significant enough to disprove them. This proves the alternative hypothesis that there is a significant change in infection, for the better, when a treatment drug is used. We can conclude that the vaccine drug is effective assuming these statistical results were calculated correctly.  

